Waveform and Passive Beamforming Design for Intelligent Reflecting Surface-Aided Wireless Information and Power Transfer

### 2020-07-03 main.tex
:pencil: System model and problem formulation

		The concept of SWIPT were first cast in \cite{Varshney2008}, where the authors investigated the Rate-Energy (R-E) tradeoff for a flat Gaussian channel and some discrete channels. Two practical receiver structures were then proposed in \cite{Zhou2013}, namely Time Switching (TS) that switches between energy harvesting (EH) and Information Decoding (ID) modes, and Power Splitting (PS) that splits the received signal into individual components. On top of this, \cite{Zhang2013} characterized the R-E region for a Multiple-Input Multiple-Output (MIMO) broadcast system under TS and PS setup. Information and power beamforming was then considered in multiuser Multi-Input Single-Output (MISO) systems to maximize the Weighted Sum-Power (WSP) subjective to Signal-to-Interference-plus-Noise Ratio (SINR) constraints \cite{Xu2014}. Motivated by this, \cite{Krikidis2014} investigated fundamental transceiver modules, information and power scheduling, and interference management for SWIPT systems. However, \cite{Boshkovska2015} pointed out that the Radio Frequency-to-Direct Current (RF-to-DC) conversion efficiency depends on the harvester input power level, which also suggested a parametric harvester model based on curve fitting and proposed an iterative resource allocation algorithm. From another perspective, \cite{Trotter2009,Boaventura2011} demonstrated that multisine waveform is more suitable for WPT as it outperforms single tone in both operation range and RF-to-DC efficiency. \cite{Clerckx2016a} derived a tractable nonlinear harvester model based on the Taylor expansion of diode I-V characteristics and proposed an adaptive waveform optimization algorithm to maximize the output DC current under rate constraints. Simulation and experiments demonstrated the benefit of modelling rectifier nonlinearity in system design \cite{Kim2019,Kim2019a}. The work was extended to SWIPT in \cite{Clerckx2018} where a superposition of modulated information waveform and multisine power waveform is optimized to enlarge the R-E region. In contrast, \cite{Park2018} suggested an adaptive dual-mode SWIPT, which alternates between single-tone transmission that exploits conventional modulation for high-rate applications and multisine transmission that encodes the information in the Peak-to-Average Ratio (PAPR) for power-demanding applications. By assuming On-Off-Keying (OOK) where bit 1 carries energy, \cite{Hu2019} compared unary and Run-Length-Limited (RLL) code in terms of rate vs battery overflow/underflow probability, and adapted conventional modulation schemes to ensure WPT is only activated at the points with large offset. Also, a learning approach \cite{Varasteh2019c} demonstrated that the offset of the power symbol is positively correlated to the harvester energy constraint, while the information symbols are symmetrically distributed around the origin. It confirmed that the superposed waveform is feasible to enlarge R-E region when considering rectifier nonlinearity. SWIPT was also explored in the network design. \cite{Liu2016} proposed a cooperative SWIPT Non-Orthogonal Multiple Access (NOMA) protocol with three user selection schemes such that the strong user assists the EH of the weak user. In \cite{Mao2019}, SWIPT based on Rate Splitting (RS) was formulated as a Weighted Sum-Rate (WSR) maximization problem subject to total harvested energy constraint for separated Information Receivers (IRs) and Energy Receivers (ERs).
	\end{subsection}

@@

\begin{section}{Single-User Optimization}
	Consider a single-user waveform and IRS optimization problem where $\boldsymbol{\alpha}=\boldsymbol{1}^{N \times 1}$. We characterize the rate-energy region through a current maximization problem subject to transmit power, rate, and IRS constraints
	\begin{maxi!}
			{\boldsymbol{w}_I,\boldsymbol{w}_P,\boldsymbol{\phi},\rho}{z(\boldsymbol{w}_I,\boldsymbol{w}_P,\boldsymbol{\phi},\rho)}{\label{op:su}}{\label{eq:su_target}}
			\addConstraint{\frac{1}{2}({\boldsymbol{w}_I^H}{\boldsymbol{w}_I}+{\boldsymbol{w}_P^H}{\boldsymbol{w}_P})\le{P}}
			\addConstraint{\sum_{n}{\log_2\left(1+\frac{(1-\rho)\lvert(h_{D,n}+\boldsymbol{v}_n^H\boldsymbol{\phi})w_{I,n}\rvert^2}{\sigma_n^2}\right)} \ge \bar{R}}\label{eq:su_rate_constraing}
			\addConstraint{\lvert{\phi_l}\rvert=1, \quad l=1,\dots,L}
			\addConstraint{0 \le \rho \le 1}
		\end{maxi!}
	Problem \ref{op:su} is intricate due to the non-convex objective function \ref{eq:su_target} and constraint \ref{eq:su_rate_constraing} with coupled variables. To reduce the design complexity, we propose an suboptimal AO algorithm that updates the IRS phase shift, transmit waveform, and receive splitting ratio iteratively.

@@

			where $\boldsymbol{M}=[\boldsymbol{V}^H,\boldsymbol{h}_D]^H \in \mathbb{C}^{(L+1) \times N}$. On top of this, we formulate the FF-IRS phase shift optimization subproblem as follows. With a given waveform $\boldsymbol{w}_I,\boldsymbol{w}_P,\rho$, introduce auxiliary variables

@@

			\begin{paragraph}{Receive Splitting Ratio}
				We then optimize the power splitting ratio $\rho$ for any fixed phase shift $\boldsymbol{\phi}$ and waveform $\boldsymbol{w}_{I/P}$. In this case, the output DC current is also expressed in \ref{eq:z_waveform} but as a function of $\rho$ with constant $t_{I/P,n}'$ given by \ref{eq:t'}. Since $z(\rho)$ is a monotonically increasing function over $\rho \in [0, 1]$, we replace the convex objective function $z(\rho)$ with affine $\rho$ and transform problem \ref{op:su} to
				\begin{maxi!}
					{\rho}{\rho}{\label{op:su_ratio}}{\label{eq:su_ratio_target}}
					\addConstraint{\sum_{n}{\log_2\left(1+\frac{(1-\rho)\lvert h_n w_n \rvert^2}{\sigma_n^2}\right)} \ge \bar{R}\label{eq:su_ratio_constraint}}
					\addConstraint{0 \le \rho \le 1}
				\end{maxi!}
				The optimal power splitting ratio $\rho^\star$ can be obtained by solving problem \ref{op:su_ratio}.
			\end{paragraph}

			\begin{paragraph}{Alternating Optimization: Inner Loop}
				As summarized in Algorithm \ref{al:inner_loop}, the inner loop updates the waveform and splitting ratio iteratively until convergence.
				\begin{algorithm}
					\caption{Inner Loop: Waveform and Splitting Ratio}
					\label{al:inner_loop}
					\begin{algorithmic}[1]
						\State \textbf{Input} $\beta_2,\beta_4,\boldsymbol{h}_D,\boldsymbol{h}_I,\boldsymbol{h}_R,P,Q,\bar{R},\epsilon,\rho,\sigma_n \ \forall n$
						\State \textbf{Initialize} $\boldsymbol{w}_{I/P}^{(0)}$
						\State $i \gets 0$
						\Repeat
						\State $i \gets i + 1$
						\State Update splitting ratio $\rho^{(i)}$ by solving problem \ref{op:su_ratio}
						\State Update waveform $\boldsymbol{w}_{I/P}^{(i)}$ by Algorithm \ref{al:waveform_sdr}
						\Until $\lvert (z^{(i)}-z^{(i-1)}) / z^{(i)} \rvert \le \epsilon$
						\State \textbf{Output} $\boldsymbol{w}_{I/P}^{\star}, \rho^{\star}$
					\end{algorithmic}
				\end{algorithm}
			\end{paragraph}
		\end{subsubsection}
	\end{subsection}

>>>>

		The concept of SWIPT were first cast in \cite{Varshney2008}, where the authors investigated the Rate-Energy (R-E) tradeoff for a flat Gaussian channel and some discrete channels. Two practical receiver structures were then proposed in \cite{Zhou2013}, namely Time Switching (TS) that switches between Energy Harvesting (EH) and Information Decoding (ID) modes, and Power Splitting (PS) that splits the received signal into individual components. On top of this, \cite{Zhang2013} characterized the R-E region for a Multiple-Input Multiple-Output (MIMO) broadcast system under TS and PS setup. Information and power beamforming was then considered in multiuser Multi-Input Single-Output (MISO) systems to maximize the Weighted Sum-Power (WSP) subjective to Signal-to-Interference-plus-Noise Ratio (SINR) constraints \cite{Xu2014}. Motivated by this, \cite{Krikidis2014} investigated fundamental transceiver modules, information and power scheduling, and interference management for SWIPT systems. However, \cite{Boshkovska2015} pointed out that the Radio Frequency-to-Direct Current (RF-to-DC) conversion efficiency depends on the harvester input power level, which also suggested a parametric harvester model based on curve fitting and proposed an iterative resource allocation algorithm. From another perspective, \cite{Trotter2009,Boaventura2011} demonstrated that multisine waveform is more suitable for WPT as it outperforms single tone in both operation range and RF-to-DC efficiency. \cite{Clerckx2016a} derived a tractable nonlinear harvester model based on the Taylor expansion of diode I-V characteristics and proposed an adaptive waveform optimization algorithm to maximize the output DC current under rate constraints. Simulation and experiments demonstrated the benefit of modelling rectifier nonlinearity in system design \cite{Kim2019,Kim2019a}. The work was extended to SWIPT in \cite{Clerckx2018} where a superposition of modulated information waveform and multisine power waveform is optimized to enlarge the R-E region. In contrast, \cite{Park2018} suggested an adaptive dual-mode SWIPT, which alternates between single-tone transmission that exploits conventional modulation for high-rate applications and multisine transmission that encodes the information in the Peak-to-Average Ratio (PAPR) for power-demanding applications. By assuming On-Off-Keying (OOK) where bit 1 carries energy, \cite{Hu2019} compared unary and Run-Length-Limited (RLL) code in terms of rate vs battery overflow/underflow probability, and adapted conventional modulation schemes to ensure WPT is only activated at the points with large offset. Also, a learning approach \cite{Varasteh2019c} demonstrated that the offset of the power symbol is positively correlated to the harvester energy constraint, while the information symbols are symmetrically distributed around the origin. It confirmed that the superposed waveform is feasible to enlarge R-E region when considering rectifier nonlinearity. SWIPT was also explored in the network design. \cite{Liu2016} proposed a cooperative SWIPT Non-Orthogonal Multiple Access (NOMA) protocol with three user selection schemes such that the strong user assists the EH of the weak user. In \cite{Mao2019}, SWIPT based on Rate Splitting (RS) was formulated as a Weighted Sum-Rate (WSR) maximization problem subject to total harvested energy constraint for separated Information Receivers (IRs) and Energy Receivers (ERs).
	\end{subsection}

@@

\begin{section}{Single-User Optimization}
	Consider a single-user waveform and IRS optimization problem where $\boldsymbol{\alpha}=\boldsymbol{1}^{N \times 1}$. We characterize the rate-energy region through a current maximization problem subject to transmit power, rate, and IRS constraints
	\begin{maxi!}
		{\boldsymbol{w}_I,\boldsymbol{w}_P,\boldsymbol{\phi},\rho}{z(\boldsymbol{w}_I,\boldsymbol{w}_P,\boldsymbol{\phi},\rho)}{\label{op:su}}{\label{eq:su_target}}
		\addConstraint{\frac{1}{2}({\boldsymbol{w}_I^H}{\boldsymbol{w}_I}+{\boldsymbol{w}_P^H}{\boldsymbol{w}_P})\le{P}}
		\addConstraint{\sum_{n}{\log_2\left(1+\frac{(1-\rho)\lvert(h_{D,n}+\boldsymbol{v}_n^H\boldsymbol{\phi})w_{I,n}\rvert^2}{\sigma_n^2}\right)} \ge \bar{R}}\label{eq:su_rate_constraint}
		\addConstraint{\lvert{\phi_l}\rvert=1, \quad l=1,\dots,L}
		\addConstraint{0 \le \rho \le 1}
	\end{maxi!}
	Problem \ref{op:su} is intricate due to the non-convex objective function \ref{eq:su_target} and constraint \ref{eq:su_rate_constraint} with coupled variables. To reduce the design complexity, we propose an suboptimal AO algorithm that updates the IRS phase shift, transmit waveform, and receive splitting ratio iteratively.

@@

			where $\boldsymbol{M}=[\boldsymbol{V}^H,\boldsymbol{h}_D]^H \in \mathbb{C}^{(L+1) \times N}$. On top of this, we formulate the FF-IRS phase shift optimization subproblem as follows. With given waveform $\boldsymbol{w}_{I/P}$ and splitting ratio $\rho$, introduce auxiliary variables

@@

			\begin{paragraph}{Receive Splitting Ratio}
				We then optimize the power splitting ratio $\rho$ for any fixed phase shift $\boldsymbol{\phi}$ and waveform $\boldsymbol{w}_{I/P}$. In this case, the output DC current is also expressed in \ref{eq:z_waveform} but as a function of $\rho$ with constant $t_{I/P,n}'$ given by \ref{eq:t'}. Since $z(\rho)$ is a quadratic function that monotonically increases over $\rho \in [0, 1]$, we replace the convex objective function $z(\rho)$ with affine $\rho$ and transform problem \ref{op:su} to
				\begin{maxi!}
					{\rho}{\rho}{\label{op:su_ratio}}{\label{eq:su_ratio_target}}
					\addConstraint{\sum_{n}{\log_2\left(1+\frac{(1-\rho)\lvert h_n w_n \rvert^2}{\sigma_n^2}\right)} \ge \bar{R}\label{eq:su_ratio_constraint}}
					\addConstraint{0 \le \rho \le 1}
				\end{maxi!}
				The optimal power splitting ratio $\rho^\star$ can be obtained by solving problem \ref{op:su_ratio}.
			\end{paragraph}

			\begin{paragraph}{Inner Loop}
				As summarized in Algorithm \ref{al:inner_loop}, the inner loop updates the waveform and splitting ratio iteratively until convergence.
				\begin{algorithm}
					\caption{Inner Loop: Waveform and Splitting Ratio}
					\label{al:inner_loop}
					\begin{algorithmic}[1]
						\State \textbf{Input} $\beta_2,\beta_4,\boldsymbol{h}_D,\boldsymbol{h}_I,\boldsymbol{h}_R,P,Q,\bar{R},\epsilon,\rho,\sigma_n \ \forall n$
						\State \textbf{Initialize} $\boldsymbol{w}_{I/P}^{(0)}$
						\State $i \gets 0$
						\Repeat
						\State $i \gets i + 1$
						\State Update splitting ratio $\rho^{(i)}$ by solving problem \ref{op:su_ratio}
						\State Update waveform $\boldsymbol{w}_{I/P}^{(i)}$ by Algorithm \ref{al:waveform_sdr}
						\Until $\lvert (z^{(i)}-z^{(i-1)}) / z^{(i)} \rvert \le \epsilon$
						\State \textbf{Output} $\boldsymbol{w}_{I/P}^{\star}, \rho^{\star}$
					\end{algorithmic}
				\end{algorithm}
			\end{paragraph}
		\end{subsubsection}
	\end{subsection}
